Fitting metabolic models to dissolved oxygen data: The Estuarine BAyesian Single-station Estimation method

### 2023-10-16 manu-draft.qmd
apacmptab now kable

```{r, results = 'hide'}
library(knitr)
library(here)
library(officer)
library(officedown)
library(flextable)
library(dplyr)

@@

Pairwise comparisons of the daily estimates of $NEM$, $P$, and $R$ returned by each method were evaluated using simple summary statistics, including Pearson correlation coefficients ($\rho$) and root mean square differences (RMSD) of the linear regression fit comparing the same flux type between methods. Results were assessed separately for estimates using the observed or detided dissolved oxygen time series. @fig-apacmpfig and `r run_reference(id = "tbl-apacmptab")` provide a comparison of the metabolic estimates from the EBASE and Odum methods. Correlations for all metabolic estimates, regardless of the input time series, were positive and significantly correlated, as expected. However, correlations were generally stronger (and RMSDs smaller) for the metabolic results based on the detided dissolved oxygen time series. The observed time series clearly produced a much larger range of values for results from both the EBASE and Odum methods.

The comparisons in @fig-apacmpfig and `r run_reference(id = "tbl-apacmptab")` suggest that tidal effects on continuous monitoring data can severely violate assumptions required for methods of estimating ecosystem metabolism [as described in detail in @Beck15]. These effects are easily identified by stronger correlations and lower RMSD using the detided dissolved oxygen. Further, effects of tidal advection can be seen with estimates from the Odum method as negative values for P and R, i.e., "anomalous" values. The effects of tidal advection on estimates from EBASE were more subtle as anomalous values cannot be returned based on constraints of the priors (i.e., none can be negative), although the range of values for the metabolic estimates were much larger with observed dissolved oxygen. A similar result (larger estimates for P and R) was observed using the synthetic time series with instrument and tidal noise (@fig-synapanoscmp). Overall, the results suggest that detiding the dissolved oxygen time series prior to estimating metabolic rates reduces tidal noise and produces more stable estimates with smaller ranges, regardless of the method used.

>>>>

```{r, results = 'hide'}
library(knitr)
library(here)
library(dplyr)

@@

Pairwise comparisons of the daily estimates of $NEM$, $P$, and $R$ returned by each method were evaluated using simple summary statistics, including Pearson correlation coefficients ($\rho$) and root mean square differences (RMSD) of the linear regression fit comparing the same flux type between methods. Results were assessed separately for estimates using the observed or detided dissolved oxygen time series. @fig-apacmpfig and @tbl-apacmptab provide a comparison of the metabolic estimates from the EBASE and Odum methods. Correlations for all metabolic estimates, regardless of the input time series, were positive and significantly correlated, as expected. However, correlations were generally stronger (and RMSDs smaller) for the metabolic results based on the detided dissolved oxygen time series. The observed time series clearly produced a much larger range of values for results from both the EBASE and Odum methods.

The comparisons in @fig-apacmpfig and @tbl-apacmptab suggest that tidal effects on continuous monitoring data can severely violate assumptions required for methods of estimating ecosystem metabolism [as described in detail in @Beck15]. These effects are easily identified by stronger correlations and lower RMSD using the detided dissolved oxygen. Further, effects of tidal advection can be seen with estimates from the Odum method as negative values for P and R, i.e., "anomalous" values. The effects of tidal advection on estimates from EBASE were more subtle as anomalous values cannot be returned based on constraints of the priors (i.e., none can be negative), although the range of values for the metabolic estimates were much larger with observed dissolved oxygen. A similar result (larger estimates for P and R) was observed using the synthetic time series with instrument and tidal noise (@fig-synapanoscmp). Overall, the results suggest that detiding the dissolved oxygen time series prior to estimating metabolic rates reduces tidal noise and produces more stable estimates with smaller ranges, regardless of the method used.
